\documentclass{tufte-handout}
\input{../common-lab}

\begin{document}
\input{../common-course}
\input{electron-eom-title}

\vspace{0.5cm}
\noindent{\bf \LARGE Week 1}\\
\vspace{0.5cm}

Hand in your responses to these questions at the end of lab today.  Please email me either an electronic file or a legible photo/scan of a handwritten page(s).  Each person should hand in their own.

\begin{enumerate}

\item Assume you don't know the dimensions of a sheet of notebook paper. Use a couple of pieces of paper to estimate
the ratio of the height of the paper to the width.

\item What is the mass of an electron? What is the charge on an electron? (you can look these values up)

\item What is the force on a charge due to an electric field?

\item What is the force on a charge due to a magnetic field?  What other factors does it depend on (other than the charge and the field)?

\item What is the change in potential energy of a charge that moves from one electric potential to another?  What determines if this is an increase or decrease in potential energy?

\item \label{q-velocity} If an electron ``falls'' from rest through a potential to accelerate it, what would its velocity squared be at the end? Derive a relationship, then calculate the velocity an electron would have after
being accelerated through a 1 volt potential. 

\item \label{q-shape} If a moving electron (say the electron from question~\ref{q-velocity}!) enters a place where there is a uniform magnetic field, what would the shape of its path be?

\item \label{q-newton} Using Newton's second law and the acceleration due to circular motion (spoiler for question~\ref{q-shape}!), derive the relationship between the electron charge, the electron mass, the velocity of the electron, the strength of the magnetic field, and the radius of the path of the electron.  I used the notation $e$, $m_e$, $v$, $B$, and $r$, respectively, for these quantities, but you are free to use any notation you like provided you use it consistently in this lab, it is understandable, and easily remembered.  

\item Derive an expression you can use to find the charge-to-mass ratio of the electron: combine the velocity squared from question~\ref{q-velocity} and the relationship from question~\ref{q-newton} to find an expression for $e/{m_e}$ in terms of the accelerating potential, the magnetic field strength, and the radius of the electrons path ($V_a$, $B$, and $r$ in my notation).  

\item Use the manual for the Teltron electron deflection tube U19155 [1000651] to draw a wiring diagram to use the tube in magnetic deflection mode.  

\item Experimentally, we will use Helmholtz coils to produce a fairly uniform magnetic field for our electron.  What is special about the geometry of Helmholtz coils?  Find (meaning lookup) the relationship that gives the magnetic field in terms of the current in the coils, properties of the coils such as radius and number of turns of wire, and fundamental constants.

\item Estimate the specs (voltage and current capabilities) of the power supply(s) you will need.

\end{enumerate}

\end{document}

 
