\documentclass{tufte-handout}
\input{../common-lab}

\begin{document}
\input{../common-course}
\input{electron-diffraction-title}

\vspace{0.5cm}
\noindent{\bf \LARGE Week 1 - Prelab}\\
\vspace{0.5cm}


\section{Objective}

The purpose of this experiment is to demonstrate the wave nature of electrons. A
beam of electrons accelerated by a potential of several kilovolts is
Bragg-diffracted by a polycrystalline graphite film, and the resulting ring
pattern is analyzed to study the relationship of the electron wavelength to its
energy.

\section{Prelab questions}

Submit on Canvas your responses to these questions at the end of lab today. Each person should hand in their own.

\begin{enumerate}

\item What is meant by the {\em phase} of a wave?  What is meant by the {\em phase difference} of two waves? What do {\em in-phase} and {\em out-of-phase} mean?  How do those ideas relate to {\em constructive} and {\em destructive} interference?

\item Consider two waves of equal wavelength $\lambda$ that originate from 
different sources but meet at the same point.
If the waves are perfectly {\it in phase}, what behavior do you expect where the waves meet?
What does it mean if they are {\it in phase}? Draw any pictures you need to explain this process.

\item Consider two waves of equal wavelength $\lambda$ and assume that they start in phase at their respective
points of origin. If one wave travels a longer distance $\Delta L$ such that $\Delta L = \lambda$
What do you expect when the waves meet? What about for the following scenarios?
\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=0pt,partopsep=0pt]
    \item $\Delta L = \frac{1}{2}\lambda$
    \item $\Delta L = \frac{3}{2}\lambda$
    \item $\Delta L = 3\lambda$
    \item $\Delta L = 978\lambda$
    \item $\Delta L = \frac{2}{5}\lambda$
    \item $\Delta L = 1.769\lambda$
\end{itemize}

\item Review the concepts of the double-slit interference pattern. Draw a sketch of this setup and
all relevant quantities that you would want to measure. How would you explain the pattern that appears on the 
viewing screen?

\item Suppose you have two narrow slits separated by distance $d$. A viewing screen is distance $L$ away. If you are
using light of wavelength $\lambda$, what is the separation between the bright spot in the middle of
the viewing screen and the first maximum on one side? If you use light of a smaller wavelength, does
the separation get larger or smaller?
\end{enumerate}


\subsection{Bragg Diffraction, Bragg's Law}

When a beam of plane electromagnetic waves (such as X-rays) of a single
wavelength strikes a crystalline solid, the incident wave interacts with the
charged particles in each atom of the solid, causing them to emit individual
secondary waves which combine to form an outgoing diffracted beam.  Bragg showed
that this process could be described by a simple physical picture.  One draws
sets of equivalent parallel planes through the atoms of the crystal; each
distinct set of planes is called a ``family'' of planes.  If the incoming beam
makes an angle $\theta$ with the planes in a given family, then the outgoing
beam makes the same angle with the planes---as if the beam were simply {\em
  reflected} from the plane.  The outgoing beam is generally rather weak, but
when the atoms in adjacent planes emit waves that are in phase, then these waves
combine constructively to produce an intense outgoing beam, known as a ``Bragg
reflection''.  These intense reflected beams occur when Bragg's Law is
satisfied, i.e., when

\begin{equation}
2 d \sin\theta = n\lambda
\label{eq:bragglaw}
\end{equation}

\noindent where $\theta$ is the angle between the incoming beam and a plane in
the family, $d$ is the perpendicular distance between adjacent planes of the
family, and $n = 1,2,3,\ldots$.\sidenote{Here, $n$ is called the \textit{order}
  of the Bragg reflection.}

As we will discover in this experiment, a beam of electrons exhibits
similar Bragg reflections to those of a beam of electromagnetic waves,
when incident on a crystal.

\begin{enumerate}
\item If you have an electron accelerated from rest across a 1 V potential difference, what is it's deBroglie wavelength? 
What about a 1000 V potential difference?

\item What is the difference between the arrangement of atoms in a crystal versus a glass or plastic?

\item In a crystal, what is the average spacing between atoms? This will vary from material to material 
so give a rough answer. {\bf Bonus:} Estimate the atomic spacing of copper using the periodic table, the density of copper, and Avogadro's number.

\item To what energy would you need to accelerate electrons such that their wavelength was equal to
what you estimate for the atomic spacings in a crystal?  What accelerating voltage would you need?

\end{enumerate}

\section{Experimental design}

Look at the equipment available to measure with (e- diffraction tube and power supply)
Now, please design a procedure for the measurements described above.\\

One you have submitted that, please read the full instructions we have provided on how to make your measurements.  COMMENT in Canvas on the agreement or difference from your own plan.

\end{document}


