\begin{itemize}[itemsep=0pt,parsep=0pt,topsep=0pt,partopsep=0pt]
\item What is an equation of motion? What is it used for and what does it give you?
\item How are Newton's laws used to get the equations of motion, specifically
Newton's Second Law.
\item How does air resistance differ from other forces? How is it parametrized?
How does it affect the equations of motion?
\item You should be able to use conservation of both linear and angular momentum
to get the equations of motion of a system. How are rocket problems different
from other simpler systems and how are those equations manipulated?
\item How do you use conservation of energy to solve for the equation
of motion of a system?
\item What does it mean for a force to be conservative? What are the requirements
for a force to be conservative? You should be able to show that a force is
conservative.
\item What concepts are involved in a simple harmonic oscillator? How is a damping or
driving force incorporated into the equations and what effect does it have
on the final equations of motion?
\item You should be able to write down the Lagrangian of a system, apply the Euler-Lagrange
equations, and get the equations of motion.
\item You should be able to write down the Hamiltonian of a system, apply Hamiltons
equations, and get the equations of motion.
\item How do you set up Newton's Second Law for a non-inertial reference frame? Where
does the centrifugal force come from? Where does the coriolis force come from? How
do they influence the motion of a body in a rotating reference frame?
\end{itemize}

